Solving \\ the \\ 15-Puzzle \\ \& \\ Rubik's Cube

### 2022-09-11 latex/Chapters/DRLSearch.tex
update

@@

Since my implementation of \textbf{DQL} is no longer just a cost-to-go estimate as with \textbf{DRL}, but a joint cost-to-go and distribution over actions (to be concrete, that means a 5-dimensional vector for the \textbf{SP} and a 13-dimensional vector for the \textbf{RC}). This leaves us with the question of how can we use the output of such a network to search for a solution. One simple answer (which I have tried, see e.g. subsection \ref{sec:S33DRLDQL}) is that we could disregard the distribution and only take into account the cost-to-go by simply using $A^{*}$. The rationale for doing so is that we might hope that jointly learning the cost-to-go and the actions via a combination of MSE and cross-entropy-loss produces a better network and leads to better heuristics, since we use more information during the learning and iterations.
\\
\\
Another possibility is to use an algorithm which combines both the actions distribution and the value function to inform its search. The Monte Carlo Tree Search (\textbf{MCTS}) is such an algorithm. In short, it generates at each step (and possibly in a distributed or multithreaded manner) paths to a \textit{leaf} on the frontier of unexpanded nodes, expands that node and checks if it is a goal state (and keeps going until the goal is found). The paths followed from initial state to leaves are decided via a combination of the probability distribution over actions (following the most promising actions) as well as the value function, according to a trade-off which can be tuned via a hyper-parameter. In order to prevent always following the same paths (which might take us very far from the goal), \textbf{MCTS} algorithms usually keep track of actions they have followed already and underweight their probability so that subsequent paths explore other actions. Let me first present some pseudo-code for the paths generation, in particular the trade-off between actions-values (exploration-exploitation in some sense):
\\

@@

\\ \teal{\algorithmiccomment{... frequent actions penalized as denominator grows faster than numerator N ... }}\black
\State{$Q \gets N . \frac{actions\_probas}{actions\_count}$}
\State{$V \gets actions\_values$}
\\ \teal{\algorithmiccomment{... high c gives weight to actions to favour exploration ... }}\black
\\ \teal{\algorithmiccomment{... return child for which cQ + V is maximized ... }}\black
\State{$action\_id \gets \Call{IndexMax}{Q + V}$}
\\
\Return{\Call{Child}{$node, action\_id$}}
\EndFunction
\end{algorithmic}
\end{algorithm}
\black

\teal
\paragraph{}{\textbf{pseudo code -- \textbf{MCTS}}}
\begin{pseudocode}
#################################################################
def MCTS(initial_node, time_out, q_v_network):
    """ pseudo code for Monte Carlo Tree Search """
    initialize_time_out(time_out)
    tree = Tree(initial_node)
    while True:
        check_time_out()
        leaf = construct_path_to_new_leaf(tree, q_v_network)
        if leaf.is_goal():
            return BFS(tree) # path to leaf can usually be improved by BFS
#################################################################
def construct_path_to_new_leaf(tree, q_v_network):
    leaf = tree.initial_node
    while leaf.expanded:
        actions_probas, actions_values = q_v_network(leaf)
        """ choose next node based on joint actions probas & values
        as well as history of actions taken
        """
        leaf = leaf.choose_child(actions_probas,
                                 actions_values)
    # add children of leaf to tree
    leaf.expand()
    # penalize path taken to favour exploration
    leaf.increment_historical_actions_count()
    return leaf
#################################################################
\end{pseudocode}
\black

>>>>

\State{\Call{InitialiseTimeOut}{$time\_out$}}

@@

Since my implementation of \textbf{DQL} is no longer just a cost-to-go estimate as with \textbf{DRL}, but a joint cost-to-go and distribution over actions (to be concrete, that means a 5-dimensional vector for the \textbf{SP} and a 13-dimensional vector for the \textbf{RC}). This leaves us with the question of how can we use the output of such a network to search for a solution. One simple answer (which I have tried, see e.g. subsection \ref{sec:S33DRLDQL}) is that we could disregard the distribution and only take into account the cost-to-go by simply using $A^{*}$. The rationale for doing so is that we might hope that jointly learning the cost-to-go and the actions via a combination of MSE and cross-entropy-loss produces a better network and leads to better heuristics, since we use more information during the learning and iterations.
\\
\\
Another possibility is to use an algorithm which combines both the actions distribution and the value function to inform its search. The Monte Carlo Tree Search (\textbf{MCTS}) is such an algorithm. In short, it generates at each step (and possibly in a distributed or multithreaded manner) paths to a \textit{leaf} on the frontier of unexpanded nodes, expands that node and checks if it is a goal state (and keeps going until the goal is found). The paths followed from initial state to leaves are decided via a combination of the probability distribution over actions (following the most promising actions) as well as the value function, according to a trade-off which can be tuned via a hyper-parameter. In order to prevent always following the same paths (which might take us very far from the goal), \textbf{MCTS} algorithms usually keep track of actions they have followed already and underweight their probability so that subsequent paths explore other actions. Let me first present some pseudo-code for the paths generation, in particular the trade-off between actions-values (exploration-exploitation):
\\

@@

\\ \teal{\algorithmiccomment{... frequent actions penalized as denominator grows faster than numerator N ... }}\black
\State{$Q \gets N . \frac{actions\_probas}{actions\_count}$}
\State{$V \gets actions\_values$}
\\ \teal{\algorithmiccomment{... high c gives weight to actions to favour exploration ... }}\black
\\ \teal{\algorithmiccomment{... return child for which cQ + V is maximized ... }}\black
\State{$action\_id \gets \Call{IndexMax}{c.Q + V}$}
\\
\Return{\Call{Child}{$node, action\_id$}}
\EndFunction
\end{algorithmic}
\end{algorithm}
\black

Then the actual \textbf{MCTS} algorithm is simply generating a new path from root to a new leaf at each iteration, until it finds the goal:

\teal
\begin{algorithm}[H]
\caption{Monte Carlo Tree Search}\label{alg:TheoryMCTS}
\begin{algorithmic}
\Function{MCTS}{$root, time\_out, q\_v\_network, c$}
\State{\Call{InitialiseTimeOut}{$time\_out$}}
\State{$tree \gets \Call{Tree}{root}$}
\If{\Call{IsGoal}{$root$}}
  \Return{$root$}
\EndIf
\While{true}
	\State\Call{CheckTimeOut}{void}
	\State{$leaf \gets root$}
	\While{$leaf.expanded$}
        		\State{$actions\_probas, actions\_values \gets q\_v\_network(leaf)$}
	           \State{$leaf \gets\Call{ChooseChildNode}{leaf,  actions\_probas, actions\_values, leaf.actions\_count, c}$}
		 \State{\Call{Expand}{$leaf, tree$}}
	\EndWhile
           \Call{IncrementActionsFromLeafToRoot}{$leaf$}
       \\ \teal{\algorithmiccomment{path to leaf can usually be improved by re-running \textbf{BFS} on the resulting tree}}\black
       \If{\Call{IsGoal}{$leaf$}}
            \Return{\Call{BlindSearch}{$tree, search\_type=Search.BFS$}}
       \EndIf
\EndWhile
\EndFunction
\end{algorithmic}
\end{algorithm}
\black
